Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
changes.

In this paper, we propose a new idea that is applicable to radial
basis functions (RBFs). Their strength is the ability to randomly
distribute points across complex physical domains, and have an
implementation that is independent of dimensionality. RBFs approximate
a function $f(\rvec)\subset \mathbb{R}^d$ sampled at a set of $N$
distinct point locations, $x_j$, by linearly combining translates of a
single radially symmetric function $\phi(r)$, where $r =
\|\rvec-\rvec_{j}\|$ denotes the Euclidean distance %(e.g., in 2-D
%$\sqrt{(x-x_j)^2+(y-y_j)^2}$) 
between where the function is evaluated
$\rvec$ and where the RBF is centered $\rvec_{j}$. That is, the
interpolant is $s(\rvec) = \sum_{j=1}^{N} w_j
\phi_i(\|\rvec-\rvec_{j}\|)$. The weights $w_j$ are obtained by
inverting the system
\begin{equation}
\parray{lccr}{
\phi_{11} & \phi_{12} & \cdots & \phi_{1N} \\
\vdots & \ddots & \vdots & \vdots \\
\phi_{n1} & \phi_{n2} & \cdots & \phi_{NN} 
}
\parray{c}{ w_{1} \\ \vdots \\ w_{N} }
=
\parray{c}{ f(\rvec_1) \\ \vdots\\ f(\rvec_N) }. 
\label{eq:rbf}
\end{equation}
where $\phi_{ij} = \phi(\|\rvec_i-\rvec_j\|)$. 
%\end{equation}
%\begin{equation}
%\begin{bmatrix}
%\phi(\|\xvec_{1}-\xvec_{1}\|) & \phi(\|\xvec_{1}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{1}-\xvec_{N}\|) \\
%\vdots & \ddots & \vdots  \\
%\phi (\|\xvec_{n}-\xvec_{1}\|) & \phi(\|\xvec_{n}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{N}-\xvec_{N}\|)
%\end{bmatrix}

@@

For a total of $N$ points, there will be $N$ linear systems to solve,
each of size $n_z \times n_z$. Each linear solve produces a row of the
RBF-FD differentiation matrix $A_{n_z}$, resulting in a $N \times N$
matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
points in the domain, one multiplies the source vector $x$ by the 
derivative matrix $A_{n_z}$ to obtain the vector $y$ of derivative values, 
$$
  y = A_{n_z} x
$$
computation of a single derivative has been reduced to a SpMV, 
where each row has $n_z$ nonzeros\footnote{We use non-conventional 
symbology for differentiation matrix, source and destination vectors
anticipating what follows.}. 
In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
three-dimensional flows. These numbers are on a par with what is used in
finite-element codes. Notice that the nonzero elements of the matrix
correspond to a relation of nearest neighbor. This relationship is not
necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
implies a non-symmetric sparsity pattern in the derivative matrix.

@@

5\todo{Add few sentence with discretization. Check Erik}
In discrete form, Equation (\ref{eq:Q}) takes the form
$$
  y^{i,j} = A^i x^j 
$$
where $j$ ranges over the number of source vectors, $i$ ranges of the number of derivative
matrices and there are $i*j$ output vectors $y$.

>>>>

In this paper, we propose a new idea that is applicable to radial
basis functions (RBFs). Their strength is the ability to randomly
distribute points across complex physical domains, and have an
implementation that is independent of dimensionality. RBFs approximate
a function $f(\rvec)\subset \mathbb{R}^d$ sampled at a set of $N$
distinct point locations, $x_j$, by linearly combining translates of a
single radially symmetric function $\phi(r)$, where $r =
\|\rvec-\rvec_{j}\|$ denotes the Euclidean distance %(e.g., in 2-D
%$\sqrt{(x-x_j)^2+(y-y_j)^2}$) 
between where the function is evaluated
$\rvec$ and where the RBF is centered at $\rvec_{j}$. That is, the
interpolant is $s(\rvec) = \sum_{j=1}^{N} w_j
\phi_i(\|\rvec-\rvec_{j}\|)$. The weights $w_j$ are obtained by
inverting the system
\begin{equation}
\parray{lccr}{
\phi_{11} & \phi_{12} & \cdots & \phi_{1N} \\
\vdots & \ddots & \vdots & \vdots \\
\phi_{n1} & \phi_{n2} & \cdots & \phi_{NN} 
}
\parray{c}{ w_{1} \\ \vdots \\ w_{N} }
=
\parray{c}{ f(\rvec_1) \\ \vdots\\ f(\rvec_N) }. 
\label{eq:rbf}
\end{equation}
where $\phi_{ij} = \phi(\|\rvec_i-\rvec_j\|)$. 
%\end{equation}
%\begin{equation}
%\begin{bmatrix}
%\phi(\|\xvec_{1}-\xvec_{1}\|) & \phi(\|\xvec_{1}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{1}-\xvec_{N}\|) \\
%\vdots & \ddots & \vdots  \\
%\phi (\|\xvec_{n}-\xvec_{1}\|) & \phi(\|\xvec_{n}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{N}-\xvec_{N}\|)
%\end{bmatrix}

@@

For a total of $N$ points, there will be $N$ linear systems to solve,
each of size $n_z \times n_z$. Each linear solve produces a row of the
RBF-FD differentiation matrix $A_{n_z}$, resulting in a $N \times N$
matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
points in the domain, one multiplies the source vector $x$ by the 
derivative matrix $A$ to obtain the vector $y$ of derivative values, 
$$
  y = A x
$$
The computation of a single derivative has been reduced to a SpMV, 
where each row has $n_z$ nonzeros\footnote{We use non-conventional 
symbology for differentiation matrix, source and destination vectors
anticipating what follows.}. 
In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
three-dimensional flows. These numbers are on a par with what is used in
finite-element codes. Notice that the nonzero elements of the matrix
correspond to a relation of nearest neighbor. This relationship is not
necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
implies a non-symmetric sparsity pattern in the derivative matrix.

@@

%\todo{Add few sentence with discretization. Check Erik}
In discrete form, Equation (\ref{eq:Q}) takes the form
$$
  y^{i,j} = A^i x^j 
$$
where $j$ ranges over the number of source vectors, $i$ ranges of the number of derivative
matrices and there are $i*j$ output vectors $y$.
